API design for machine learning software: experiences from the scikit-learn project

### 2013-09-11 paper.tex
some more trimming

To illustrate the initialize-fit sequence,
let us consider a supervised learning task using logistic regression.
Given the API defined above, solving this problem is as simple as the following
example.
\begin{pythoncode}
from sklearn.linear_model import LogisticRegression

clf = LogisticRegression(penalty="l1")
clf.fit(X_train, y_train)
\end{pythoncode}
In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
setting the \texttt{penalty} hyper-parameter to \texttt{"l1"} for
$\ell_1$ regularization. Other hyper-parameters (such as \texttt{C},
the strength of the regularization) are not explicitly given and
thus set to the default values. Upon calling \texttt{fit}, a model is
learned from the training arrays \texttt{X\_train} and \texttt{y\_train},
and stored on the object for later use.
%% (In this example, \texttt{fit} is called solely for its side effects.)
Since all estimators share the same interface, using a different learning algorithm is
as simple as replacing the constructor (the class name);
to build a random forest on
the same data, one would simply replace
\texttt{LogisticRegression(penalty="l1")} in the snippet above by
\texttt{RandomForestClassifier()}.

@@

The \textit{predictor} interface extends the notion of an estimator
by adding a \texttt{predict}
method that takes an array \texttt{X\_test} and produces
predictions for \texttt{X\_test}, based on the learned parameters of the
estimator (we call the input to \texttt{predict} ``\texttt{X\_test}'' in order
to emphasize that \texttt{predict} generalizes to new data). In the case of
supervised learning estimators, this method typically returns the predicted
labels or values computed by the model.  Continuing with the previous example,
predicted labels for \texttt{X\_test} can be obtained using the following
snippet:
\begin{pythoncode}
y_pred = clf.predict(X_test)
\end{pythoncode}

Some unsupervised learning estimators may also implement the \texttt{predict}
interface. The code in the snippet below fits a $k$-means model with $k=10$ on
training data \texttt{X\_train}, and then uses the  \texttt{predict} method to
obtain cluster labels (integer indices) for unseen data \texttt{X\_test}.
\begin{pythoncode}
from sklearn.cluster import KMeans

km = KMeans(n_clusters=10)
km.fit(X_train)
clust_pred = km.predict(X_test)
\end{pythoncode}

@@

Finally, predictors must provide a \texttt{score} function to assess their
performance on a batch of input data. In supervised estimators, this method
takes as input arrays \texttt{X\_test} and \texttt{y\_test} and typically
computes the coefficient of determination between \texttt{y\_test} and
\texttt{predict(X\_test)} in regression, or the accuracy
in classification.
The only requirement is that the \texttt{score} method return a value
that quantifies the quality of its predictions (the higher, the better).
An unsupervised estimator may also expose a \texttt{score} function
to compute, for instance, the likelihood of the given data under its model.

@@

Since it is common to modify or filter data before feeding it to a learning
algorithm, some estimators in the library implement a \textit{transformer}
interface which defines a \texttt{transform} method. It takes as input some new
data \texttt{X\_test} and yields as output a transformed version of
\texttt{X\_test}. Preprocessing, feature selection, feature extraction and dimensionality reduction
algorithms are all provided as transformers.  In our example,
to standardize the input \texttt{X\_train} to zero mean and unit variance
before fitting the logistic regression estimator,
one would write:
\begin{pythoncode}
from sklearn.preprocessing import StandardScaler

scaler = StandardScaler()
scaler.fit(X_train)
X_train = scaler.transform(X_train)
\end{pythoncode}
Of course, in practice, it is important to apply the same preprocessing to the
test data \texttt{X\_test}. Since a \texttt{StandardScaler} estimator stores the
mean and standard deviation that it computed for the training set, transforming
an unseen test set \texttt{X\_test} maps it into the appropriate region of
feature space:
\begin{pythoncode}
X_test = scaler.transform(X_test)
\end{pythoncode}
Transformers also include a variety of learning algorithms, such as
dimensionality reduction (PCA, manifold learning), kernel approximation,
and other mappings from one feature space to another.

Every transformer allows \texttt{fit(X).transform(X)}
to be written as \texttt{fit\_transform(X)}.
The combined \texttt{fit\_transform} method prevents repeated computations.
Depending on the transformer,
it may skip only an input validation step,
or in fact use a more efficient algorithm for the transformation.
In the same spirit, clustering algorithms provide a
\texttt{fit\_predict} method
that is equivalent to \texttt{fit} followed by \texttt{predict},
returning cluster labels assigned to the training samples.

@@

Some machine learning algorithms are expressed naturally
as meta-algorithms parametrized on simpler algorithms.
Examples include ensemble methods which
build and combine several simpler models (boosting, bagging, random forests),
or multiclass classification schemes which combine multiple binary classifiers.
In \sklearn, such algorithms are implemented as \textit{meta-estimators}.
They take as input an estimator which is used
internally for learning and making predictions.
All meta-estimators implement the regular estimator interface.
% XXX bagging example here

Since meta-estimators require users to construct nested objects,
the decision to implement a meta-estimator
rather than integrate the behavior it implements
into existing estimator classes
is always based on a trade-off between generality and ease of use.

@@

In \sklearn, model selection is supported in two distinct meta-estimators,
\texttt{GridSearchCV} and \texttt{RandomizedSearchCV}.  They take as input an
estimator (basic or composite), whose hyper-parameters must be optimized, and a
set of hyperparameter settings to search through.
This set is represented as a mapping of parameter names
to a set of discrete choices in the case of grid search,
which exhaustively enumerates the ``grid'' (cartesian product)
of complete parameter combinations.
Randomized search is a smarter algorithm
that avoids the combinatorial explosion in grid search
by sampling a fixed number of times from its parameter distributions
(see \citealp{bergstra2012}).

@@

param_grid = [
  {"kernel": ["linear"], "C": [1, 10, 100, 1000]},
  {"kernel": ["rbf"], "C": [1, 10, 100, 1000],
   "gamma": [0.001, 0.0001]}
]

clf = GridSearchCV(SVC(), param_grid, scoring="f1", cv=10)
clf.fit(X_train, y_train)
\end{pythoncode}
\caption{Optimizing the hyper-parameters of an SVM for $F_1$ score.}
\label{svm-gridsearch}
\end{figure*}

>>>>

To illustrate the initialize-fit sequence,
let us consider a supervised learning task using logistic regression (LR).
Given the API defined above, solving this problem is as simple as the following
example.
\begin{pythoncode}
from sklearn.linear_model import LogisticRegression
clf = LogisticRegression(penalty="l1")
clf.fit(X_train, y_train)
\end{pythoncode}
In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
setting the \texttt{penalty} hyper-parameter to \texttt{"l1"} for
$\ell_1$ regularization. Other hyper-parameters (such as \texttt{C},
the strength of the regularization) are not explicitly given and
thus set to the default values. Upon calling \texttt{fit}, a model is
learned from the training arrays \texttt{X\_train} and \texttt{y\_train},
and stored on the object for later use.
%% (In this example, \texttt{fit} is called solely for its side effects.)
Since all estimators share the same interface, using a different learning algorithm is
as simple as replacing the constructor (the class name);
to build a random forest on
the same data, one would simply replace
\texttt{LogisticRegression(penalty="l1")} in the snippet above by
\texttt{RandomForestClassifier()}.

@@

The \textit{predictor} interface extends the notion of an estimator
by adding a \texttt{predict}
method that takes an array \texttt{X\_test} and produces
predictions for \texttt{X\_test}, based on the learned parameters of the
estimator (we call the input to \texttt{predict} ``\texttt{X\_test}'' in order
to emphasize that \texttt{predict} generalizes to new data). In the case of
supervised learning estimators, this method returns the predicted
labels or values computed by the model:
\begin{pythoncode}
y_pred = clf.predict(X_test)
\end{pythoncode}

Some unsupervised learning estimators may also implement the \texttt{predict}
interface. The code below fits a $k$-means model with $k=10$ on
training data \texttt{X\_train}, and then uses the  \texttt{predict} method to
obtain cluster labels for unseen data \texttt{X\_test}.
\begin{pythoncode}
from sklearn.cluster import KMeans
km = KMeans(n_clusters=10)
km.fit(X_train)
clust_pred = km.predict(X_test)
\end{pythoncode}

@@

Finally, supervised predictors provide a \texttt{score} function to assess their
performance on a batch of input data. This method
takes as input arrays \texttt{X\_test} and \texttt{y\_test} and typically
computes the coefficient of determination between \texttt{y\_test} and
\texttt{predict(X\_test)} in regression, or the accuracy
in classification.
The only requirement is that the \texttt{score} method return a value
that quantifies the quality of its predictions (the higher, the better).
An unsupervised estimator may also expose a \texttt{score} function
to compute, for instance, data likelihood under its model.

@@

Since it is common to modify or filter data before feeding it to a learning
algorithm, some estimators in the library implement a \textit{transformer}
interface which defines a \texttt{transform} method. It takes as input some new
data \texttt{X\_test} and yields as output a transformed version of
\texttt{X\_test}. Preprocessing, feature selection, feature extraction and dimensionality reduction
algorithms are all provided as transformers.  In our example,
to standardize the input \texttt{X\_train} to zero mean and unit variance
before fitting the LR estimator, one would write:
\begin{pythoncode}
from sklearn.preprocessing import StandardScaler
scaler = StandardScaler()
scaler.fit(X_train)
X_train = scaler.transform(X_train)
\end{pythoncode}
Of course, in practice, it is important to apply the same preprocessing to the
test data \texttt{X\_test}. Since a \texttt{StandardScaler} estimator stores the
mean and standard deviation that it computed for the training set, transforming
an unseen test set \texttt{X\_test} maps it into the appropriate region of
feature space:
\begin{pythoncode}
X_test = scaler.transform(X_test)
\end{pythoncode}
Transformers also include a variety of learning algorithms, such as
dimensionality reduction (PCA, manifold learning), kernel approximation,
and other mappings from one feature space to another.

Every transformer allows \texttt{fit(X).transform(X)}
to be written as \texttt{fit\_transform(X)}.
This combined operation prevents repeated computations,
either by skipping only an input validation step,
or by using a more efficient algorithm for the transformation.
In the same spirit, clustering algorithms provide a \texttt{fit\_predict}
that is equivalent to \texttt{fit} followed by \texttt{predict},
returning cluster labels for the training samples.

@@

Some learning algorithms are expressed naturally
as meta-algorithms parametrized on simpler ones.
Examples include ensemble methods which
build and combine several simpler models (boosting, bagging, random forests),
or multiclass classification schemes which combine multiple binary classifiers.
In \sklearn, such algorithms are implemented as \textit{meta-estimators}.
They take as input an estimator which is used
internally for learning and making predictions.
All meta-estimators implement the regular estimator interface.
% XXX bagging example here

Since meta-estimators require users to construct nested objects,
the decision to implement a meta-estimator
rather than integrate its functionality
into existing classes
is always based on a trade-off between generality and ease of use.

@@

In \sklearn, model selection is supported by the meta-estimators
\texttt{GridSearchCV} and \texttt{RandomizedSearchCV}.  These take as input an
estimator (basic or composite), whose hyper-parameters must be optimized, and a
set of hyperparameter settings to search through.
This set is represented as a mapping of parameter names
to a set of discrete choices in the case of grid search,
which exhaustively enumerates the ``grid'' (cartesian product)
of complete parameter combinations.
Randomized search is a smarter algorithm
that avoids the combinatorial explosion in grid search
by sampling a fixed number of times from its parameter distributions
\citep{bergstra2012}.

@@

param_grid = [
  {"kernel": ["linear"], "C": [1, 10, 100, 1000]},
  {"kernel": ["rbf"], "C": [1, 10, 100, 1000],
   "gamma": [0.001, 0.0001]}
]
clf = GridSearchCV(SVC(), param_grid, scoring="f1", cv=10)
clf.fit(X_train, y_train)
\end{pythoncode}
\caption{Optimizing the hyper-parameters of an SVM for $F_1$ score.}
\label{svm-gridsearch}
\end{figure*}
